Implementing Real-Time Transport Services over an Ossified Network

### 2016-05-23 paper.tex
Fixes typos

\documentclass[10pt]{sig-alternate-05-2015}
\usepackage[l2tabu,orthodox]{nag}
\usepackage[utf8x]{inputenc}
\usepackage[british]{babel}
\usepackage[babel=true]{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage{balance}
\usepackage{tabularx}
\usepackage{rotating}
\usepackage{tikz}
\usepackage{xspace}   %MF
\usepackage{enumitem} %MF

@@

@@

Real-time applications require a set of transport services not currently
provided by widely-deployed transport protocols. Ossification prevents the
deployment of a novel protocols, restricting solutions to protocols using
either TCP or UDP as a substrate. We describe the transport services
required by real-time applications. We show that, in the short term (i.e.,
while UDP is blocked at current levels), TCP offers a feasible substrate
for providing these services. Over the longer term, protocols using UDP
may reduce the number of networks blocking UDP, enabling a shift towards
its use as a demultiplexing layer for novel transport protocols.

@@

In this paper we identify and present the appropriate set of transport
services and APIs for real-time applications, and demonstrate their
merit by implementing a proof-of-concept. We show that it is possible
to realise real-time services and APIs in the context of both TCP and
UDP, despite the limitations imposed by their legacies, and by
middleboxes, and ossification of the network.  Initial experiments
with our implementation suggest that
%to show that, despite network ossification,
the network has the flexibility to deploy new transport protocols, provided care
is taken to reinterpret application and transport layer boundaries that respect
conventional UDP and TCP.
%and deploy those new transports in the context of TCP and UDP.

@@

% % When compared to conventional real-time systems, networked multimedia
% % deadlines are neither soft nor hard.
% Networked multimedia deadlines are unusual when compared to other real-time
% systems. They are simultaneously flexible and strict: flexible in that the exact
% value of the deadline is typically not important, provided it is of the right
% order-of-magnitude for the application, but strict in that any particular
% deadline provides a hard cut-off, after which the data is useless.
It is useful to distinguish between networked multimedia deadlines and
deadlines in conventional real-time systems. A multimedia deadline is neither
soft, since arrivals played after a deadline contribute to poor experience; nor
are deadlines hard, since occasional missed messages can be tolerated.

\subsection{Partial Reliability}
In a best-effort network, deadlines
constrain packet delivery service to \emph{partial reliability}.
%The network has non-zero probability of losing any particular packet.
For example, when used to repair loss, the limits of forward error correction
imply some probability that packet will be non-recoverable. By contrast,
retransmissions used to recover from loss have potentially unbounded delay
(since any retransmission may itself be lost).
% These probabilities can be estimated, to give probabilistic constraints on
% loss and delay, but will be non-zero. Accordingly, to meet deadlines the
Accordingly, a transport protocol that meets deadlines should provide partial
reliability, acknowledging that it may be unable to deliver all data by its
deadline.

@@

In the context of both deadlines and dependencies coupled with packet loss, partial reliability
requires application-level framing \cite{clark:1990:architecture} to
make the best use of payload data. At the transport
layer, this implies a \emph{message oriented} service, that
maintains application data unit (ADU) boundaries.

@@

We note that a connection-oriented transport is a lesser, but useful,
requirement.
% for many real-time multimedia applications.
Indeed, flexibility to
change the destination within a call is beneficial for
applications that support mobile users, or for some forms of multiparty session.
However, to support NAT traversal and to help dynamically manage firewall
pinholes, it is often desirable for the transport to be connection oriented. We
believe these concerns outweigh the benefits of connectionless transport, and so
add a requirement for connection oriented service. Similarly, while not strictly
needed by the applications, it is beneficial if the transport provides a
keep-alive service to refresh NAT and firewall bindings if the application goes
silent.

@@

\begin{itemize}[nosep]
  \item Hosts setup and tear-down sockets using the \texttt{socket()}
    and \texttt{close()} functions, as in the standard Berkeley sockets
    API.

@@

    It is instructive to compare the partially reliable send and receive
    functions to their TCP-like counterparts. The \texttt{send\_message()} call
    takes four additional parameters compared to the standard Berkeley Sockets
    API. These are 1) a message sequence number, that can be used to re-order
    messages and detect message loss; 2) a relative deadline, which is combined
    with an estimate of the current round-trip-time, and the time that the
    message has spent in the sending buffer, to determine if a message will
    arrive in time to be played-out; 3) the message sequence number of any
    message on which this depends, for example, of a video I-frame on which a
    P-frame is predicted; and 4) a sub-stream identifier, used, for example, to
    differentiate audio, video, sub-title, control, and repair streams.

@@

\begin{table*}
  \small
  \centering
  \begin{tabularx}{\textwidth}{llll}
    \toprule
      Transport Service            & Function & Parameters                                     & Return Value \\
    \midrule
                                   & \texttt{socket}
                                              & \texttt{af} -- Address family                  & Socket descriptor \\
                                   &          & \texttt{st} -- Socket type & \\
                                   & \texttt{close}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
    \midrule
      Connection oriented          & \texttt{bind}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   &          & \texttt{addr} -- Address to bind to & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{listen}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   & \texttt{accept}
                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
                                   &          & \texttt{addr} -- Address of peer & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{connect}
                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
    \midrule
      Deadlines                    & \texttt{set\_po\_delay}
                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
    \midrule
      Message oriented             & \texttt{send\_message}
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
                                   &          & \texttt{buf} -- Message data & \\
                                   &          & \texttt{len} -- Length of message data & \\
                                   &          & \texttt{seq\_num} -- Sequence number & \\
      Deadlines                    &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\
      Dependencies                 &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
                                   & \texttt{recv\_message}
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
                                   &          & \texttt{buf} -- Buffer for message data        & Substream identifier \\
                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
    \bottomrule
  \end{tabularx}
  \caption{Outline transport API for real-time applications. Return values
  shown are for successful calls; in all cases, -1 is returned in the event
  of an error}
  \label{tab:api}
\end{table*}

@@

The TCP congestion control algorithm is executed by the end points, and
can be changed, provided the new version requires no new information to
be exchanged.
%\todo{expand, and discuss how this evolution has occurred to date}
We note that, while standardised TCP congestion control has followed the
goal of maximising throughput at the expense of latency and variability,
this is not required by the protocol. TCP Vegas \cite{brakmo:1994:tcp-vegas}
is perhaps the best known approach that changes these constraints, with a
delay-based algorithm that reduces latency. It would also be possible to
implement alternatives that seek stability, or compatibility with the
dictates of a video codec, rather than traditional ``TCP Friendly''
congestion control -- even if implemented within TCP.

@@

The standard transport protocols, TCP and UDP, are not well-suited for
real-time applications. Both can be made to work, but the existence of
numerous papers exploring how to make media play-out over TCP reliable,
and almost as extensive a collection discussing UDP-based protocol design,
suggests that this is difficult to do well. To make effective use of the
network, and simplify real-time application design and implementation, we
need to deploy new transport services and protocols that allow innovative
applications to be developed by users who are not experts in transport
protocol design. We discussed requirements such for a new transport, in
the context of the TAPS framework, and outlined a straw-man abstract API,
in Section \ref{sec:services}.

It seems likely that the right long-term approach for doing this is to
repurpose UDP as a demultiplexing layer for higher-layer protocols. We can
then deploy an appropriate transport protocol framework as a user-space
library, that can be reused as appropriate. In the short term, however,
there are sufficient networks that block UDP, that any new transport
protocol needs to be able to run over TCP. Sections \ref{sec:ossification}
and \ref{sec:partial} discuss how this can be done, and suggest from some
initial measurement studies that this may be feasible to deploy. Section
\ref{sec:realising} considers prototypes that present such services over
UDP, and presents our initial prototype demonstrated for TCP-based use.

>>>>

\documentclass[10pt]{sig-alternate-05-2015}
\usepackage[l2tabu,orthodox]{nag}
\usepackage[utf8x]{inputenc}
\usepackage[british]{babel}
\usepackage[babel=true]{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage{balance}
\usepackage{tabularx}
\usepackage{rotating}
\usepackage{tikz}

@@

\usepackage{xspace}

@@

Real-time applications require a set of transport services not currently
provided by widely-deployed transport protocols. Ossification prevents the
deployment of novel protocols, restricting solutions to protocols using
either TCP or UDP as a substrate. We describe the transport services
required by real-time applications. We show that, in the short-term (i.e.,
while UDP is blocked at current levels), TCP offers a feasible substrate
for providing these services. Over the longer term, protocols using UDP
may reduce the number of networks blocking UDP, enabling a shift towards
its use as a demultiplexing layer for novel transport protocols.

@@

In this paper we identify and present the appropriate set of transport
services and APIs for real-time applications, and demonstrate their
merit by implementing a proof-of-concept. We show that it is possible
to realise real-time services and APIs in the context of both TCP and
UDP, despite the limitations imposed by their legacies, and by
middleboxes, and ossification of the network.  Initial experiments
with our implementation suggest that
%to show that, despite network ossification,
the network has the flexibility to deploy new transport protocols,
provided care is taken to reinterpret layer boundaries that respect
conventional UDP and TCP.\todo{I can infer the meaning of previous sentence having worked on the problem; otherwise may be very ambiguous to the reader.}
%and deploy those new transports in the context of TCP and UDP.

@@

Networked multimedia deadlines are unusual when compared to other real-time
systems. They are simultaneously flexible and strict: flexible in that the exact
value of the deadline is typically not important, provided it is of the right
order-of-magnitude for the application, but strict in that any particular
deadline provides a hard cut-off, after which the data is useless. \todo{Is this para necessary at all? Its contribution is unclear and/or unconvincing -- flexible is fine; strict is less believable, \ie
how are RT systems any different when the deadline expires? (Many RTSs will
discard, too. Ultimately, I get what it is trying to say; I'm less sure it needs saying) Maybe the following(?):}
% When compared to conventional real-time systems, networked multimedia
% deadlines are neither soft nor hard.
It is useful to draw a distinction between in networked multimedia deadlines and
deadlines in conventional real-time systems. A multimedia deadline is neither soft, since arrivals played after a deadline contribute to poor experience; nor are deadlines hard, since occasional missed messages can be tolerated.

\subsection{Partial Reliability}
In a network that offers only a best-effort packet delivery service, deadlines
constrain delivery to \emph{partial reliability}.
%The network has non-zero probability of losing any particular packet.
For example, when used to repair loss, the limits of forward error correction
imply some probability that packet will be non-recoverable. By contrast,
retransmissions used to recover from loss have potentially unbounded delay
(since any retransmission may itself be lost).
% These probabilities can be estimated, to give probabilistic constraints on
% loss and delay, but will be non-zero. Accordingly, to meet deadlines the
Accordingly, a transport protocol that meets deadlines should provide partial
reliability, acknowledging that it may be unable to deliver all data by its
deadline.

@@

Partial reliability in the context of both deadlines and dependencies
requires application-level framing \cite{clark:1990:architecture} to
make the best use of payload data, irrespective of packet loss. At the transport
layer, this implies that we provide a \emph{message oriented} service, that
maintains application data unit (ADU) boundaries.

@@

We note that a connection-oriented transport is a lesser, but useful,
requirement.
% for many real-time multimedia applications.
Indeed, flexibility to
change the destination of data within a call could be beneficial for
applications that support mobile users, or for some forms of multiparty session.
However, to support NAT traversal and to help dynamically manage firewall
pinholes, it is often desirable for the transport to be connection oriented. We
believe these concerns outweigh the benefits of connectionless transport, and so
add a requirement for connection oriented service. Similarly, while not strictly
needed by the applications, it is beneficial if the transport provides a
keep-alive service to refresh NAT and firewall bindings if the application goes
silent.

@@

\begin{itemize}
  \item Hosts setup and tear-down sockets using the \texttt{socket()}
    and \texttt{close()} functions, as in the standard Berkeley sockets
    API.

@@

    The \texttt{send\_message()} call takes four additional parameters
    compared to the standard Berkeley Sockets API. These are 1) a
    message sequence number, that can be used to re-order messages and
    detect message loss; 2) a relative deadline, which is combined
    with an estimate of the current round-trip-time, and the time that
    the message has spent in the sending buffer, to determine if a
    message will arrive in time to be played-out; 3) the message
    sequence number of any message on which this depends, for example,
    of a video I-frame on which a P-frame is predicted; and 4) a
    sub-stream identifier, used, for example, to differentiate audio,
    video, sub-title, control, and repair streams.

@@

\begin{table*}
  \small
  \centering
  \begin{tabularx}{\textwidth}{llll}
    \toprule
      Transport Service            & Function & Parameters                                     & Return Value(s) \\
    \midrule
                                   & \texttt{socket}
                                              & \texttt{af} -- Address family                  & Socket descriptor \\
                                   &          & \texttt{st} -- Socket type & \\
                                   & \texttt{close}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
    \midrule
      Connection oriented          & \texttt{bind}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   &          & \texttt{addr} -- Address to bind to & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{listen}
                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
                                   & \texttt{accept}
                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
                                   &          & \texttt{addr} -- Address of peer & \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
                                   & \texttt{connect}
                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
    \midrule
      Deadlines                    & \texttt{set\_po\_delay}
                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
    \midrule
      Message oriented             & \texttt{send\_message}
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
                                   &          & \texttt{buf} -- Message data & \\
                                   &          & \texttt{len} -- Length of message data & \\
                                   &          & \texttt{seq\_num} -- Sequence number & \\
      Deadlines                    &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\
      Dependencies                 &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
                                   & \texttt{recv\_message}
                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
                                   &          & \texttt{buf} -- Buffer for message data        & Substream identifier \\
                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
    \bottomrule
  \end{tabularx}
  \caption{Outline transport API for real-time applications. Return values
  shown are for successful calls; in all cases, -1 is returned in the event
  of an error}
  \label{tab:api}
\end{table*}

@@

The TCP congestion control algorithm is executed by the end points, and
can be changed, provided the new version requires no new information to
be exchanged.
\todo{expand, and discuss how this evolution has occurred to date}
We note that, while standardised TCP congestion control has followed the
goal of maximising throughput at the expense of latency and variability,
this is not required by the protocol. TCP Vegas \cite{brakmo:1994:tcp-vegas}
is perhaps the best known approach that changes these constraints, with a
delay-based algorithm that reduces latency. It would also be possible to
implement alternatives that seek stability, or compatibility with the
dictates of a video codec, rather than traditional ``TCP Friendly''
congestion control -- even if implemented within TCP.

@@

The standard transport protocols, TCP and UDP, are not well-suited for
real-time applications. Both can be made to work, but the existence of
numerous papers exploring how to make media play-out over TCP reliable, 
and almost as extensive a collection discussing UDP-based protocol design,
suggests that this is difficult to do well. To make effective use of the
network, and simplify real-time application design and implementation, we
need to deploy new transport services and protocols that allow innovative
applications to be developed by users who are not experts in transport 
protocol design. We discussed requirements for such a new transport, in
the context of the TAPS framework, and outlined a straw-man abstract API,
in Section \ref{sec:services}.

It seems likely that the right long-term approach for doing this is to
repurpose UDP as a demultiplexing layer for higher-layer protocols. We can
then deploy an appropriate transport protocol framework as a user-space
library, that can be reused as appropriate. In the short-term, however, 
there are sufficient networks that block UDP, that any new transport
protocol needs to be able to run over TCP. Sections \ref{sec:ossification}
and \ref{sec:partial} discuss how this can be done, and suggest from some
initial measurement studies that this may be feasible to deploy. Section
\ref{sec:realising} considers prototypes that present such services over
UDP, and presents our initial prototype demonstrated for TCP-based use.
